---
title: "Lecture 2: CRS - Homework"
author: Tyler Hoecker
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

## Setup 
Load general packages, set filepath for data files.
```{r}
library (tidyverse)
library(rgdal)
data_path = "/Users/tylerhoecker/GitHub/Zoo955_fork/Data/"
```


## 1) Download the LTER lake shapefiles from the LTER database. Map Lake Mendota. Add a point for the location of the CFL.

### Using ```library(sp)``` and ```library(raster)```
Load packages for this section.
```{r}
library(sp)
library(raster)
```

Download North Temperatre Lakes LTER study lakes spatial data from a URL and save it to 'Data' directory.
```{r eval = FALSE}
download_link <- 'https://lter.limnology.wisc.edu/file/11669/download?token=tWhRExjmwpMK9caKoMd2VcgKmFln3dKHVTC0mwUBvqg'
temp <- tempfile()
download.file(download_link,temp)
unzip(temp, exdir = paste0(data_path,'NTL_LTER_Yahara_Lakes'))
unlink(temp)
```

Load shapefile.
```{r}
yahara_lakes <- readOGR(dsn = paste0(data_path,'NTL_LTER_Yahara_Lakes'), layer = "yld_study_lakes")
```

Extract just Lake Mendota and transform CRS to WGS84.
```{r}
lake_mendota <- 
  yahara_lakes[yahara_lakes$SHAIDNAME == 'Lake Mendota',] %>% 
  spTransform(., CRS("+init=epsg:4326 +proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))
```

Create a point for the CFL (looked up using Google Maps), and define its CRS to be the same as ```lake_mendota```
```{r}
cfl_pt <- 
  data.frame(x = -89.402923, y = 43.077269) %>% 
  SpatialPoints() %>% 
  `crs<-` (crs(lake_mendota))
```

Plot the lake shapefile and the CFL point
``` {r}
plot(gridlines(lake_mendota), lty=2)
plot(lake_mendota, col = 'lightblue2', add=TRUE)
plot(cfl_pt, pch = 20, col = 'red', cex = 2, add = TRUE)
```

### Or, do the same thing using ```library(sf)```
I like the sf package because it plays nicely with the tidyverse, and treats spatial objects more like regular dataframes.

Load packages for this section.
```{r}
library(sf)
```

Read in ```yld_study_lakes.shp```, transform CRS to WGS84, and filter out only Lake Mendota.
```{r}
lake_mendota_sf <- 
  st_read(dsn = paste0(data_path,'NTL_LTER_Yahara_Lakes/yld_study_lakes.shp')) %>% 
  st_transform(crs = c("+init=epsg:4326 +proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")) %>%
  filter(SHAIDNAME == 'Lake Mendota')
```

Make a new sf object for the CFL. 
```{r}
cfl_sf <- data.frame(x = -89.402923, y = 43.077269) %>% 
  st_as_sf(coords = c("x", "y"), crs = st_crs(lake_mendota_sf))
```

Plot them both. 
```{r}
ggplot() +
  geom_sf(data = lake_mendota_sf, fill = 'lightblue2') +
  geom_sf(data = cfl_sf, colour = "red", size = 4) +
  labs(x = 'Longitude', y = 'Latitude') +
  theme_bw()
```

Or, plot Lake Mendota, and use ```annotate``` to add the CFL "by hand".
```{r}
ggplot() +
  geom_sf(data = lake_mendota_sf, fill = 'lightblue2') +
  annotate("point", x = -89.402923, y = 43.077269, colour = "red", size = 4) +
  labs(x = 'Longitude', y = 'Latitude') +
  theme_bw()
```




## 2) Download the National Land Cover (NLCD) dataset for 2011. Load it into R. What is the CRS?

Download NLCD dataset from URL.
```{r eval = FALSE}
download_link <- 'http://www.landfire.gov/bulk/downloadfile.php?TYPE=nlcd2001v2&FNAME=nlcd_2001_landcover_2011_edition_2014_10_10.zip'

temp <- tempfile()
#download.file(download_link,temp)
unzip(temp, exdir = 'Data/')
unlink(temp)
```

Load NLCD dataset, and view CRS.
```{r message = TRUE}
nlcd <- raster(paste0(data_path,'NLCD/nlcd_2001_landcover_2011_edition_2014_10_10.img'))

crs(nlcd)
```
# 3) What’s the best way to check that CRS of two objects are identical?

The simplest way I can think of is to compare the proj4strings as character strings. If they are completely identical, the boolean comparison will be true: ```as.character( crs(object1) ) == as.character( crs(object2) )```.

Here are some functions that could be handy:  

For sf objects:
```{r message = TRUE}
crsTest <- function(object1, object2){
  if(as.character(crs(object1)) == as.character(crs(object2))){
    print("You're in luck, the CRS's are identical!")
} else { 
    print("CRS's not equal, try again.")}
}

crsTest(lake_mendota,nlcd)
```

For sf objects:
```{r message = TRUE}
st_crsTest <- function(object1, object2){
  if(as.character(st_crs(object1)[2]) == as.character(st_crs(object2)[2])){
    print("You're in luck, the CRS's are identical!")
} else { 
    print("CRS's not equal, try again.")}
}

st_crsTest(lake_mendota_sf,cfl_sf)
```

### Turns out there's already a function for this in sp... but it's messages are pretty boring :)
```{r message = TRUE}
identicalCRS(lake_mendota,nlcd)
```
